Discrepancy, Coresets, and Sketches in Machine Learning

### 2019-05-09 karninliberty19.tex
camera ready v1

\section{Introduction}
The study of coresets in optimization as a whole and in machine learning specifically has a long history. 
The basic setup is as follows.
Suppose you are trying to optimize an expression over a set of items, data points, or examples.
The optimization problem is difficult and often square, cubic or even exponential in the input set size.
Therefore, there is a strong incentive to reduce the cardinality of that set. 
The goal is, therefore, to pinpoint a small subset of the data items which approximates the entire input set with respect to some function or geometric measure. Such small sets are called coresets. 
This idea is very general and applies to geometric properties of the data \cite{agarwal2005geometric}, 
clustering \cite{DBLP:conf/compgeom/Har-PeledK05} \cite{DBLP:conf/stoc/FeldmanL11}, classification \cite{DBLP:conf/ijcai/Har-PeledRZ07}, regression \cite{DBLP:conf/nips/MunteanuSSW18} 
machine learning \cite{bachem2017practical}, density estimation \cite{DBLP:journals/corr/abs-1802-01751}, and many other problems.

@@

Rademacher complexity (see for example \cite{Bartlett:2003:RGC:944919.944944}) is a standard measure of generalization.
In other words, bounding the Rademacher complexity is a good way to upper bound the sample complexity. 
A sample is an instance of a coreset which is chosen i.i.d.\ from the data (or the underlying distribution).
A carefully selected coreset can, at least potentially, be better than a uniformly sampled one. 
It can be smaller and still give the same generalization power or give better generalization with the same number of data points. 
There are papers such as \cite{langberg2010universal, tolochinsky2018coresets} and references therein that tie the coreset size to the VC dimension of the function family, and the average sensitivity of the dataset. 
These relationships come up as tools for constructing coresets rather than complexity measures aimed to characterize generalization ability. 
This paper defines the analog to Rademacher complexity that aims to characterize the coreset complexity, i.e., the generalization ability of the best possible coreset of a fixed size.

We show that our result applies to any analytic function of the dot product. 
These include Logistic Regression $F(q) =  \sum_i \log(1+\exp(\ip{y_ix_i,q}))$, Covariance or matrix approximation 
$F(q) =  \sum_i \ip{x_i,q}^2$, sigmoid activation loss $F(q) =  \sum_i 1/(1+\exp(\ip{y_ix_i,q}))$, linear regression $F(q) =  \sum_i (\ip{y_ix_i,q} -y_i)^2$ and many others. 
For all the aforementioned  $x,q \in R^d$ and $y_i \in \{-1,1\}$.
By bounding the class discrepancy of all such functions we obtain coresets of size $O(\sqrt{d}/\eps)$ for all of them.

We note that while we obtain a universal {\it additive guarantee} it is often much harder to get a  {\it multiplicative guarantee}.
For Logistic regression, for example, a recent paper \cite{DBLP:journals/corr/abs-1805-08571} provides a coreset with a  {\it multiplicative guarantee} that is based on an average sensitivity property of the dataset. 
They provide a lower bound for the size of a multiplicative error coreset proving in particular that in general, it is not possible to achieve $m \ll n$. 
The coreset they build is of cardinality $m \approx \mu\sqrt{nd^3}/\eps^2$ where $\mu \geq 1$ is a complexity measure of the dataset.  \cite{tolochinsky2018coresets} gives a generic multiplicative coreset construction for any monotonic function with $\ell_2^2$ regularization. 
The dependence they get is $\tilde O(d/\eps^2)$ ignoring logarithmic factors.

We show that our result also applies to any analytic function of the squared distance.
A prime example of that is Gaussian kernel density estimation.
Kernel density estimation is a popular tool in data analysis aimed to estimate a continuous distribution with a finite set of points. 
Among other applications, this tool is used for outlier detection \cite{schubert2014generalized}, regression \cite{fan2018local}, and clustering \cite{rinaldo2010generalized}. 
A thorough survey could be found in \cite{silverman2018density}. 
Given a set of $n$ data points $\{x_1,\ldots,x_n\}$ and a query $q$, the Gaussian density estimate at point $q$ is given by $\sum_i K(x_i,q) = \sum_i e^{-\|x-q\|^2}$. 
Obtaining the smallest possible coreset for this problem has been open for several years.
The state-of-the-art is given by \cite{DBLP:journals/corr/abs-1802-01751} (see references within). 
They achieve coresets of size $O(\sqrt{d\log(1/\eps)}/\eps)$ where $d$ is the dimension of the original data points. 
Their result holds for any Lipchitz bounded positive semi-definite kernels. 
The result is constructive though it is polynomial rather than (quasi-)linear in the data size. 
The authors give an almost matching lower bound of $\sqrt{d}/\eps$ and pose an open question for closing the gap between the bounds. 
Additionally, they provide an algorithm based on the Frank-Wolf method that achieves a $1/\eps^2$ sized coreset. 
It is, in fact, known (see \cite{lopez2015towards}, Theorem 1) that a uniform random sample of $\log(1/\delta)/\eps^2$ points gives a coreset w.p. $1-\delta$ for some kernel types. 
In this paper we resolve the open question by \cite{DBLP:journals/corr/abs-1802-01751} and prove that the coreset complexity of Gaussian kernel density is indeed $O(\sqrt{d}/\eps)$, matching the lower bound. 
In fact, we show that this is the coreset complexity for any analytic function of the squared distance $f(x,q) = f(\|x-q\|^2)$.

We also provide another independent result; 
a very simple and deterministic algorithm for constructing coresets of size $O(1/\eps^2)$ for any positive semi-definite kernel. 
Although the worst-case lower bound is $\Omega(1/\eps^2)$ which matches the coreset achieved by sampling, we conjecture that in realistic scenarios, the deterministic algorithm will outperform random sampling as it explicitly takes advantage of negative correlations.

@@

\begin{definition} [\bf Coreset Complexity] 
For a function $f:\X,\Q\rightarrow \R$ let $F(q) = \sum_{i=1}^{m} f(x_i,q)$ for a any set $\{x_1,\ldots,x_m\} \subset \X$.
For a set $S \subset [m]$ let $\tilde F(q) = \sum_{i \in S}w_i f(x_i,q)$ for some $w\in\R_+^m$ which is independent of $q$.
The coreset complexity of $f$ is the size of the smallest set $S$ such that $\forall q \in \Q \; |F(q)  - \tilde F(q)| \le \eps m$.
\end{definition}
\noindent The following facts are true for the common cases where $D_m = O(c/m)$ or $D_m = O(c/\sqrt{m})$. Although they were previously known (see e.g., \cite{phillips2009small}, Theorem 1.1) we give their proof here for completeness.
\begin{fact}\label{fct:eps1}
Any function $f$ with class discrepancy $D_m = O(c/m)$ has coreset complexity of $O(c/\eps)$.
\end{fact}
\begin{fact}\label{fct:eps2}
Any function $f$ with class discrepancy $D_m = O(c/\sqrt{m})$ has coreset complexity $O(c^2/\eps^2)$.
\end{fact}
\begin{proof}
\noindent For a set of $n$ points $x_1,\ldots,x_n$ and arbitrary query $q$, consider the signed-sum error function $E(q) = \sum_{i=1}^{n} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Recalling $F(q) = \sum f(x_i,q)$, we consider $\tilde F_{+}(q) = F(q) + E(q)   = \sum_{i | \sigma_i=1} 2 f(x_i,q)$ and similarly $\tilde F_{-}(q) = F(q) - E(q)  =  \sum_{i | \sigma_i=-1} 2 f(x_i,q)$. 
We have that both $\tilde F_{+}(q)$ and $\tilde F_{-}(q)$ are approximations for $F(q)$ obtained by coresets of item-weight $2$. 
The error is at most $|\tilde F_{\pm}(q)- F(q)| =  |E(q)|$, and one of the coresets are of cardinality of at most $n/2$. 
The above is true for any choice of signs $\sigma$, specifically, for those minimizing $\max_q | E(q)|$.
By definition we can select signs such that $|E(q)| \le m D_n$.

Naturally, one could iterate this process. Starting with $n$ items and ending with $k$.
Let $F_t$ denote the (unweighted) sum of functions $f$ after $t$ iterations and $n_t$ denote the cardinality of the coreset\footnote{We use the term $a=b \pm c$ to mention $|a-b| \leq c$ }. 
\[
F = F_0 = 2F_1 \pm n_0D_{n_0} = 4F_{2} \pm 2n_1D_{n_1} \pm n_0D_{n_0} = \ldots = 2^T F_T \pm \sum_{t=0}^{T-1} 2^t n_t D_{n_t}
\]
Here $T$ stand for the total number of iterations. Let us analyze the error term.
Given $n_t \le n/2^{t}$ and the polynomial dependence of $D_m$ on $m$ we have 
$$\sum_{t=0}^{T-1} 2^t n_t D_{n_t} \le n \sum_{t=0}^{T-1} D_{n/2^t} = n\cdot O(D_{n_T}) . $$ 
Setting $n_T$ in a way that $D_{n_T} = O(\eps)$ leads to a coreset with item weights of $w^T$ of the appropriate cardinality and completes the proof.

%
%\noindent  Consider the signed-sum error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
%Now consider $\tilde F_{+}(q) = F(q) + E(q)   = \sum_{i | \sigma_i=1} 2 f(x_i,q)$ and similarly $\tilde F_{-}(q) = F(q) - E(q)  =  \sum_{i | \sigma_i=-1} 2 f(x_i,q)$. 
%We have that both $\tilde F_{+}(q)$ and $\tilde F_{-}(q)$ are approximations for $F(q)$ obtained by coresets of item-weight $2$. 
%The error is at most $|\tilde F_{\pm}(q)- F(q)| =  |E(q)|$. 
%The above is true for any choice of signs $\sigma$, specifically, for those minimizing $\max_q | E(q)|$.
%By definition we can select signs such that $|E(q)| \le m D_m$. 
%
%\add{
%Naturally, one could iterate this process. Starting with $n$ items and ending with $m$.
%Let $F_t$ denote the (unweighted) sum of functions $f$ after $t$ iterations and $n_t$ denote the cardinality of the coreset. 
%\[
%F = F_0 = 2F_1 \pm n_0D_{n_0} = 4F_{2} \pm 2n_1D_{n_1} \pm n_0D_{n_0} = \ldots = 2^T F_T \pm \sum_{t=0}^{T-1} 2^t n_t D_{n_t}
%\]
%Here $T$ stand for the total number of iterations needed for reaching a coreset of size $m$. 
%Given $n_t \le n/2^{t}$ and the polynomial dependence of $D_m$ on $m$ we have 
%$\sum_{t=0}^{T-1} 2^t n_t D_{n_t} \le n \sum_{t=0}^{T-1} D_{n/2^t} = n\cdot O(D_{n_T})$ for all $T$.
%Specifically, for $T$ such that $n_T \approx m$, setting the weights $w = 2^T$ we have $F(q) = \tilde F(q) \pm O(nD_m)$.
%Substituting the above expressions for $D_m$ completes the proof.
%}
%\st{
%Naturally, one could iterate this process.
%Noting that the last term in above is a geometric sum we can conclude that 
%At step one, create a coreset of size at most $m/2$ of items of weight $2$ and incur an error of at most $c$ or $c \sqrt{m}$.
%Then, create a coreset of size $m/4$ of items of weight $4$. Here, you incur error of $2c$ or $2c\sqrt{m/2}  = c\sqrt{2m}$.
%Note that the sum of errors is a geometric sequence that is asymptotically dominated by its last element. 
%Halting the compression at $\Theta(c/\eps)$ or $\Theta(c^2/\eps^2)$ items respectively achieves the goal.
%}
\end{proof}

@@

The following it a straight forward fact which is provided mainly for convenience. 
It loosely says that optimizing models on coresets generalizes. In other words, ERM works as expected.
\begin{fact}
Let $f(x,q)$ be the loss suffered by model $q$ on example $x$. 
Let $R(q) = \frac{1}{n}F(q) = \frac{1}{n} \sum_{i=1}^n f(x_i,q)$ be the {\it empirical} risk associated with it.
Let $q^*$ denote the best empirical risk minimizer on the data ($q^* = \arg\min_q F(q)$).
Let $\tilde q$ be the minimizer of $q$ over an optimal weighted coreset of size $m$ ($\tilde q = \arg\min_q \tilde F(q)$).
Then $R(\tilde q) \le R(q^*) + O(D_m)$.
\end{fact}
\begin{proof}
This fact follows from the standard argument about empirical risk minimization.
\[
R(\tilde q) = \frac1n F(\tilde q)  \le  \frac{1}{n} \tilde F(\tilde q) + O(D_m) \le \frac{1}{n} \tilde F(q^*) + O(D_m) \le  \frac{1}{n}  F(q^*) + O(D_m)  = R(q^*)+ O(D_m)
\]
The first and last transitions are by definition. The second and fourth are by the approximation bounds above.
The third transition is due to the optimality of $\tilde q$ for $\tilde F$
\end{proof}

@@

\subsection{Towards an Efficient Algorithm}

%\section{A Simple Algorithm for Kernel Density Estimation}
From section \ref{sec:analytic} we know that the class discrepancy of the Gaussian kernel is $D_m = O(\sqrt{d}/m)$. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. 
In fact, we show that for any positive kernel $D_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section for high dimensions $d > m$. 
More importantly, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = \{x_1,\ldots, x_n\}$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as $ F(q) = \sum_{i=1}^{n} K(x_i,q) $.
Here, $K$ is any \emph{positive semi-definite kernel} function. The most frequent examples include
$$ K(x,q) = \exp(- \|x-q\|_2^2/\lambda^2),\;\; K(x,q) = \exp(- \|x-q\|/\lambda) ,\;\;  K(x,q) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on the distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

@@

\paragraph{Note} Theorem~\ref{thm:disc simple kernel} provides an upper bound of $\sqrt{m}$ for the sign discrepancy. 
This upper bound is tight since there exist sets of vectors in high dimensions that requires it. 
For data that lends itself to density estimation, however, one should expect input vectors to be clustered together.
In such cases, the algorithm above performs much better than the worst-case bound predicts. 
We leave it to future work to define properties of the data that ensure algorithm better guarantees by Algorithm~\ref{kernelAlg}.
\bibliography{density}

>>>>

\paragraph{Acknowledgments:} The authors sincerely thank Nikhil Bansal, Nikhil Srivastava, Jeff Phillips, Wai Ming Tai, and Camron Musco 
for generously sharing their time and ideas.
They helped us uncover the usefulness of Banaszczyk's theorem for proving Lemma~\ref{uc}, compare to other results on coresets and discrepancy
(specifically on Gaussian Kernel Density estimation), and understand the connection to graph sparsification and matrix column subset selection results.

\section{Introduction}
The study of coresets in optimization as a whole and in machine learning specifically has a long history. 
The basic setup is as follows.
Suppose you are trying to optimize an expression over a set of items, data points, or examples.
The optimization problem is difficult. 
Its running time dependence on the input set size is square, cubic, or even exponential.
As a result, there is a strong incentive to reduce the cardinality of that set. 
The goal is, therefore, to pinpoint a small subset of data items which approximates the entire input set with respect to the optimization at hand. 
Such small sets are called coresets. 
This idea is very general and applies to geometric properties of the data \cite{agarwal2005geometric}, 
clustering \cite{DBLP:conf/compgeom/Har-PeledK05} \cite{DBLP:conf/stoc/FeldmanL11}, classification \cite{DBLP:conf/ijcai/Har-PeledRZ07}, regression \cite{DBLP:conf/nips/MunteanuSSW18} 
machine learning \cite{bachem2017practical}, density estimation \cite{DBLP:journals/corr/abs-1802-01751}, and many other problems.

@@

Rademacher complexity (see for example \cite{Bartlett:2003:RGC:944919.944944}) is a standard measure of generalization.
In other words, bounding the Rademacher complexity is a good way to upper bound the sample complexity. 
A sample is an instance of a coreset which is chosen i.i.d.\ from the data (or the underlying distribution).
A carefully selected coreset can, at least potentially, be better than a uniformly sampled one. 
It can be smaller and still give the same generalization power or give better generalization with the same number of data points. 
There are papers such as \cite{langberg2010universal, tolochinsky2018coresets} and references therein that tie the coreset size to the VC dimension of the function family and the average sensitivity of the dataset. 
These relationships come up as tools for constructing coresets rather than complexity measures aimed to characterize generalization ability. 
This paper defines the analog to Rademacher complexity that aims to characterize the coreset complexity, i.e., the generalization ability of the best possible coreset of a fixed size.

We show that our result applies to any analytic function of the dot product. 
These include Logistic Regression $F(q) =  \sum_i \log(1+\exp(\ip{y_ix_i,q}))$, Covariance or matrix approximation 
$F(q) =  \sum_i \ip{x_i,q}^2$, sigmoid activation loss $F(q) =  \sum_i 1/(1+\exp(\ip{y_ix_i,q}))$, linear regression $F(q) =  \sum_i (\ip{y_ix_i,q} -y_i)^2$ and many others. 
For all the aforementioned  $x,q \in R^d$ and $y_i \in \{-1,1\}$.
By bounding the class discrepancy of all such functions we prove the existence of coresets of size $O(\sqrt{d}/\eps)$ for all of them.

We note that while we obtain a universal {\it additive guarantee} it is often much harder to get a  {\it multiplicative guarantee}.
For Logistic regression, for example, a recent paper \cite{DBLP:journals/corr/abs-1805-08571} provides a coreset with a  {\it multiplicative guarantee} that is based on an average sensitivity property of the dataset. 
They provide a lower bound for the size of a multiplicative error coreset proving in particular that in general, it is not possible to achieve $m \ll n$. 
The coreset they build is of cardinality $m \approx \mu\sqrt{nd^3}/\eps^2$ where $\mu \geq 1$ is a complexity measure of the dataset.  
\cite{tolochinsky2018coresets} give a generic multiplicative coreset construction for any monotonic function with $\ell_2^2$ regularization. 
The dependence they get is $\tilde O(d/\eps^2)$ ignoring logarithmic factors.

We show that our result also applies to any analytic function of the squared distance.
A prime example of that is Gaussian kernel density estimation.
Kernel density estimation is a popular tool in data analysis aimed to estimate a continuous distribution with a finite set of points. 
Among other applications, this tool is used for outlier detection \cite{schubert2014generalized}, regression \cite{fan2018local}, and clustering \cite{rinaldo2010generalized}. 
A thorough survey could be found in \cite{silverman2018density}. 
Given a set of $n$ data points $\{x_1,\ldots,x_n\}$ and a query $q$, the Gaussian density estimate at point $q$ is given by $\sum_i K(x_i,q) = \sum_i e^{-\|x-q\|^2}$. 
Obtaining the smallest possible coreset for this problem has been open for several years.
The state-of-the-art is given by \cite{DBLP:journals/corr/abs-1802-01751} (see references within). 
They achieve coresets of size $O(\sqrt{d\log(1/\eps)}/\eps)$ where $d$ is the dimension of the original data points. 
Their result holds for any Lipchitz bounded positive semi-definite kernels. 
The result is constructive though it is polynomial rather than (quasi-)linear in the data size. 
The authors give an almost matching lower bound of $\sqrt{d}/\eps$ and pose an open question for closing the gap between the bounds. 
In this paper we resolve the open question by \cite{DBLP:journals/corr/abs-1802-01751} and prove that the coreset complexity of Gaussian kernel density is indeed $O(\sqrt{d}/\eps)$, matching the lower bound. 
In fact, we show that this is the coreset complexity for any bounded analytic function of the squared distance $f(x,q) = f(\|x-q\|^2)$.

In high dimensions $\sqrt{d}/\eps$ could be large. 
It is known (see \cite{lopez2015towards}, Theorem 1) that a uniform random sample of $\log(1/\delta)/\eps^2$ points gives a coreset w.p. $1-\delta$ for some kernel types. 
\cite{DBLP:journals/corr/abs-1802-01751} provide an algorithm based on the Frank-Wolf method that achieves a $1/\eps^2$ sized coreset. 
In section \ref{kde2} we provide (as a stand alone result) a very simple and deterministic algorithm for constructing coresets of size $1/\eps^2$ for any positive semi-definite kernel.  The worst-case coreset lower bound is $\Omega(1/\eps^2)$ which matches the coreset achieved by sampling. 
Yet, for real data the deterministic algorithm outperforms random sampling significantly (experiments not included in this manuscript).

@@

\begin{definition} [\bf Coreset Complexity] 
For a function $f:\X,\Q\rightarrow \R$ let $F(q) = \sum_{i=1}^{m} f(x_i,q)$ for a any set $\{x_1,\ldots,x_m\} \subset \X$.
For a set $S \subset [m]$ let $\tilde F(q) = \sum_{i \in S}w_i f(x_i,q)$ for some $w\in\R_+^m$ which is independent of $q$.
The coreset complexity of $f$ is the size of the smallest set $S$ such that $\forall q \in \Q \; |F(q)  - \tilde F(q)| \le \eps m$.
\end{definition}
\noindent The following facts are true for the common cases where $D_m = O(c/m)$ or $D_m = O(c/\sqrt{m})$. Although they were previously known (see e.g., \cite{phillips2009small}, Theorem 1.1) we give their proof here for completeness.
\begin{fact}\label{fct:eps1}
Any function $f$ with class discrepancy $D_m = O(c/m)$ has coreset complexity of $O(c/\eps)$.
\end{fact}
\begin{fact}\label{fct:eps2}
Any function $f$ with class discrepancy $D_m = O(c/\sqrt{m})$ has coreset complexity $O(c^2/\eps^2)$.
\end{fact}
\begin{proof}
\noindent  Consider the signed-sum error function $E(q) = \sum_{i=1}^{m} \sigma_i f(x_i,q)$ where $\sigma_i \in \{-1,1\}$.
Now consider $\tilde F_{+}(q) = F(q) + E(q)   = \sum_{i | \sigma_i=1} 2 f(x_i,q)$ and similarly $\tilde F_{-}(q) = F(q) - E(q)  =  \sum_{i | \sigma_i=-1} 2 f(x_i,q)$. 
We have that both $\tilde F_{+}(q)$ and $\tilde F_{-}(q)$ are approximations for $F(q)$ obtained by coresets of item-weight $2$. 
The error is at most $|\tilde F_{\pm}(q)- F(q)| =  |E(q)|$. 
The above is true for any choice of signs $\sigma$, specifically, for those minimizing $\max_q | E(q)|$.
By Definition we can select signs such that $|E(q)| \le m D_m$.

\add{
Naturally, one could iterate this process. Starting with $n$ items and ending with $m$.
Let $F_t$ denote the (unweighted) sum of functions $f$ after $t$ iterations and $n_t$ denote the cardinality of the coreset. 
\[
F = F_0 = 2F_1 \pm n_0D_{n_0} = 4F_{2} \pm 2n_1D_{n_1} \pm n_0D_{n_0} = \ldots = 2^T F_T \pm \sum_{t=0}^{T-1} 2^t n_t D_{n_t}
\]
Here $T$ stand for the total number of iterations needed for reaching a coreset of size $m$. 
Given $n_t \le n/2^{t}$ and the polynomial dependence of $D_m$ on $m$ we have 
$\sum_{t=0}^{T-1} 2^t n_t D_{n_t} \le n \sum_{t=0}^{T-1} D_{n/2^t} = n\cdot O(D_{n_T})$ for all $T$.
Specifically, for $T$ such that $n_T \approx m$, setting the weights $w = 2^T$ we have $F(q) = \tilde F(q) \pm O(nD_m)$.
Substituting the above expressions for $D_m$ completes the proof.
}
\st{
Naturally, one could iterate this process.
Noting that the last term in above is a geometric sum we can conclude that 
At step one, create a coreset of size at most $m/2$ of items of weight $2$ and incur an error of at most $c$ or $c \sqrt{m}$.
Then, create a coreset of size $m/4$ of items of weight $4$. Here, you incur error of $2c$ or $2c\sqrt{m/2}  = c\sqrt{2m}$.
Note that the sum of errors is a geometric sequence that is asymptotically dominated by its last element. 
Halting the compression at $\Theta(c/\eps)$ or $\Theta(c^2/\eps^2)$ items respectively achieves the goal.
}
\end{proof}

@@

\add{
The following it a straight forward fact which is provided mainly for convenience. 
It loosely says that optimizing models on coresets generalizes. In other words, ERM works as expected.
\begin{fact}
Let $f(x,q)$ be the loss suffered by model $q$ on example $x$. 
Let $R(q) = \frac{1}{n}F(q) = \frac{1}{n} \sum_{i=1}^n f(x_i,q)$ be the {\it empirical} risk associated with it.
Let $q^*$ denote the best empirical risk minimizer on the data ($q^* = \arg\min_q F(q)$).
Let $\tilde q$ be the minimizer of $q$ over an optimal weighted coreset of size $m$ ($\tilde q = \arg\min_q \tilde F(q)$).
Then $R(\tilde q) \le R(q^*) + O(D_m)$.
\end{fact}
\begin{proof}
This fact follows from the standard argument about empirical risk minimization.
\[
R(\tilde q) = \frac1n F(\tilde q)  \le  \frac{1}{n} \tilde F(\tilde q) + O(D_m) \le \frac{1}{n} \tilde F(q^*) + O(D_m) \le  \frac{1}{n}  F(q^*) + O(D_m)  = R(q^*)+ O(D_m)
\]
The first and last transitions are by definition. The second and fourth are by the approximation bounds above.
The third transition is due to the optimality of $\tilde q$ for $\tilde F$
\end{proof}
}

@@

\section{A Simple Algorithm for Kernel Density Estimation}\label{kde2}
From section \ref{sec:analytic} we know that the class discrepancy of the Gaussian kernel is $D_m = O(\sqrt{d}/m)$. 
Here, we provide a computationally efficient bound that can be achieved with a straightforward algorithm of complexity $O(m^2)$. Together with the results of Section~\ref{sec:sketch} this provides an efficient sketching algorithm for Kernel Density Estimation. 
In fact, we show that for any positive kernel $D_m = O(1/\sqrt{m})$. This bound is superior to that of the previous section for high dimensions $d > m$. 
More importantly, there is a very simple, intuitive, and deterministic algorithm for computing the signs $\sigma$. 
Given a collection of data points $X = \{x_1,\ldots, x_n\}$ in $\R^d$ the density function $f: \R^d \rightarrow \R$ of a point $q$ is defined as $ F(q) = \sum_{i=1}^{n} K(x_i,q) $.
Here, $K$ is any \emph{positive semi-definite kernel} function. The most frequent examples include
$$ K(x,q) = \exp(- \|x-q\|_2^2/\lambda^2),\;\; K(x,q) = \exp(- \|x-q\|/\lambda) ,\;\;  K(x,q) = (1+\|x-q\|/_2^2/\lambda^2)^{-1}$$
where $\lambda$ is a scaling parameter. For simplicity, we assume that $K(x,x) \leq 1$ for all data points. Notice that for any kernel based on the distance we have $K(x,x)=1$ exactly for all $x \in \R^d$.

@@

\paragraph{Note} Theorem~\ref{thm:disc simple kernel} provides an upper bound of $\sqrt{m}$ for the sign discrepancy. 
This upper bound is tight since there exist sets of vectors in high dimensions that requires it. 
For data that lends itself to density estimation, however, one should expect input vectors to be clustered together.
In such cases, the algorithm above performs much better than the worst-case bound predicts. 
We leave it to future work to define properties of the data that ensure better guarantees for Algorithm~\ref{kernelAlg}.
\bibliography{karninliberty19}
